Measuring causal influence with\\ back-to-back regression: the linear case

### 2019-09-24 paper/introduction.tex
wip paper

First, causal and non-causal factors may be numerous and collinear. In physics, for example, one may be challenged to identify whether fusion is caused by a change in temperature or a change in pressure, as these two factors may, at first, be difficult to manipulate independently. This issue becomes increasingly pronounced as the number of potential factors increases. In neuroscience, for example, identifying whether the frequency of a word presented on a subject's retina modulates brain activity can be surprisingly difficult. Indeed, word frequency covaries with other factors such as their length (short words are more frequent than long words), their categories (determinants are more frequent than adverbs), their individual letters (e.g. words that contain `z' are rare) and so forth \citep{kutas2011thirty,pegado2014timing}. Instead of selecting a set of words that control for all of these factors simultaneously, it is thus common to use forward modeling, i.e. to train a mathematical model to predict observations (e.g. brain activity) from a minimal combination of competing factors (e.g. word length, word frequency, e.g. \citep{huth2016natural}), and investigate, in the model, the estimated contribution of each factor \citep{friston1994statistical}.

The second challenge to measuring causal influence is that observations can be large and complex. The relationship between causes and effects is thus often considered in a backward manner, by training models to maximally predict causes from multidimensional observations. For example, brain activity is often recorded with hundreds or thousands of sensors simultaneously. It is thus common to use `decoding' techniques, by, for example, fitting a support vector machine across multiple sensors to predict the category of a stimulus \citep{cichy2014resolving,  kriegeskorte2008representational, norman2006beyond}.

Both forward and backward modeling have competing benefits and drawbacks. Specifically, forward modeling disentangles the independent contribution of collinear factors, but does not combine multidimensional observations. By contrast, backward modeling combines multiple observations, but does not disentangle collinear factors \cite{weichwald2015causal, hebart2018deconstructing, king2018encoding}. There also exist cross-decomposition techniques such as Partial Least Squares (PLS) and Canonical Correlation Analysis (CCA), but their resulting coefficients are notoriously difficult to interpret \citep{lebart1995statistique}.

We introduce the `back-to-back regression', which combines the benefits of forward and backward modeling (Section~\ref{sec:algorithm}). After detailing the method and proving its convergence (Section~\ref{sec:theorem}), we show with synthetic data that it outperforms state-of-the-art forward, backward and cross-decomposition techniques, both in terms of estimating i) causal influence and ii) out-of-environment predictions (Section~\ref{sec:experiment_synthetic}). Finally, we apply back-to-back regression to a large neuroimaging dataset and reveal that distinct but collinear word features lead to distinguishable brain responses (Section~\ref{sec:experiment_real}).

>>>>

First, causal and non-causal factors may be numerous and collinear. In physics, for example, one may be challenged to identify whether fusion is caused by a change in temperature or a change in pressure, as these two factors may, at first, be difficult to manipulate independently. This issue becomes increasingly pronounced as the number of potential factors increases. In neuroscience, for example, identifying whether the frequency of a word presented on a subject's retina modulates brain activity can be surprisingly difficult. Indeed, the frequency of words in natural language covaries with other factors such as their length (short words are more frequent than long words), their categories (determinants are more frequent than adverbs) and so forth \citep{kutas2011thirty,pegado2014timing}. Instead of selecting a set of words that control for all of these factors simultaneously, it is thus common to use forward modeling, i.e. to train a model to predict observations (e.g. brain activity) from a minimal combination of competing factors (e.g. word length, word frequency, e.g. \citep{huth2016natural}), and investigate, in the model, the estimated contribution of each factor \citep{friston1994statistical}.

The second challenge to measuring causal influence is that observations can be large and complex. The relationship between causes and effects is thus often considered in a backward manner, by training models to maximally predict causes from multidimensional observations. For example, brain activity is often recorded with hundreds or thousands of sensors simultaneously. As multiple sensors may be affected by common noise sources, it is common to use `decoding' techniques, by, for example, fitting a support vector machine across multiple sensors to predict the category of a stimulus \citep{cichy2014resolving,  kriegeskorte2008representational, norman2006beyond}.

Both forward and backward modeling have competing benefits and drawbacks. Specifically, forward modeling disentangles the independent contribution of collinear factors, but does not combine multidimensional observations. By contrast, backward modeling combines multiple observations, but does not disentangle collinear factors \cite{weichwald2015causal, hebart2018deconstructing, king2018encoding}. To combine the benefits of forward and backward modeling, several authors have proposed to use cross-decomposition techniques such as Partial Least Squares (PLS) and Canonical Correlation Analysis (CCA) \citep{dechevigne2018CCA}.
CCA and PLS aim to find, from two sets of data $X$ and $Y$, the components $v$ and $w$ were $Xw$ and $Yv$ are maximially correlated or maximally covarying respectively. Because CCA and PLS are are based on a generalized eigen decomposition, their resulting coefficients are mixing the features of $X$ and $Y$ in a way that makes them notoriously difficult to interpret \citep{lebart1995statistique}.

Here, we introduce the `back-to-back regression', which not only combines the benefits of forward and backward modeling (Section~\ref{sec:algorithm}), but also provide a fast, robust and interpretable causal coefficients. After detailing the method and proving its convergence (Section~\ref{sec:theorem}), we show with synthetic data that it outperforms state-of-the-art forward, backward and cross-decomposition techniques in identifying causal influence (Section~\ref{sec:experiment_synthetic}). Finally, we apply back-to-back regression to a large neuroimaging dataset and reveal that distinct but collinear word features lead to distinguishable brain responses (Section~\ref{sec:experiment_real}).

### 2019-09-24 paper/meg_exps.tex
wip paper

\subsubsection{MEG preprocessing and feature generation}

The raw MEG data was bandpass-filtered between 0.1 and 40Hz using MNE-Python
default parameters \citep{gramfort2013meg, gramfort2014mne}. Specifically, we used a zero-phase finite impulse
response filter (FIR) with a Hamming window and with transition bands of 0.1Hz
and 10Hz for the low and high cut-off frequencies. The raw data was then segmented 100ms before word onset and 1s after
word onset ($t=0$ms corresponds to word onset). Finally, each resulting
segment was baseline-corrected between -100ms and 0ms, and decimated by 5 and
thus led a sampling frequency of 240Hz. Figure 1 in the supplementary material
shows the average response of the magnetometers evoked by the words presented
to the first subject. For each subject and each time sample relative to word onset, we
build an observation matrix $Y \in \mathbb{R}^{n \times d_y}$ of $n\approx$ 2700 words
by $d_y=301$ MEG channels (273 magnetometers and 28 compensation channels). Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

We annotated each word with 38 features: 26
correspond to the count of each letter in the word (from 'a' to 'z',
independently of their case), the total number of letters ('word length'), the
frequency of the word in natural language (using the Zipf logarithmic scale
of \citep{van2014subtlex}), and the part-of-speech tags (i.e. NOUN, PROPN (proper
noun), VERB, PRON (pronoun), CONJ (conjunction), DET (determinant), ADJ
(adjective), ADV (adverb), and NUM (numerical)) using Spacy \citep{spacy2} and
WordFreq \citep{speerwordfreq}. For clarity purposes, we refer to these
features using two broad categories: visual (i.e. each letter and the total
number of letters) and lexical (part-of-speech and word frequency). This
procedure yields an $X \in \mathbb{R}^{n \times d_x}$ matrix of $n\approx$ 2700 words by
$d_x=38$ features for each subject. Each of the columns of $X$ is normalized to
have a mean and a standard deviation of 0 and 1 respectively.

@@

Our goal is to identify which word features ($X$) lead to a reliable modulation of brain activity as measured with MEG recordings ($Y$) at each time sample and for each subject separately. To this aim,
we compare three methods: forward regression, backward regression, and back-to-back regression.
Corresponding results can be found in Figure \ref{fig:meg_twocurves}.

@@

Forward regression was performed with an $\ell_2$-regularized least-squares regression $H \in \mathbb{R}^{d_x \times d_y}$ from $X$ to $Y$, and estimates the causal influence $E_{i,i}$ as $\| H_{i, :} \|^2$.
Backward regression was performed with an $\ell_2$-regularized least-squares regression $G \in \mathbb{R}^{d_y \times d_x}$ from $Y$ to $X$, and estimates the causal influence $E_{i, i}$ as the Pearson correlation between the $i$-th column of $X$ and the $i$-th column of $YG$, using a validation sample. Finally, we used the back-to-back regression was performed using two $\ell_2$-regularized
least squares regressions $G$ and $H$ as described in section \ref{sec:algorithm}, using a bagging factor of 10, and alpha cross-validated between $10^{-5}$ and $10^{5}$.

The three types of analyses were implemented for each subject and each time
sample independently. To estimate the reliability of these estimates in the
population, we tested, using a one-sample $t$-test whether the distribution of
the estimated contribution of each factor differed from baseline ($t=0$ms),
as $\text{t-test}(\hat E_{t=0} - \hat E_{\tau})$, with $\tau$ ranging from
150ms to 1s after word onset. We use the resulting t and p-values as proxy to
quantify the performance of each model. These results are depicted in Figure
3 of the Supplementary Material. For concision purposes, we only report this metric averaged
between 150ms and 1s after word onset.

>>>>

\subsubsection{MEG preprocessing}

The raw MEG data was bandpass-filtered between 0.1 and 40Hz using MNE-Python
default parameters \citep{gramfort2013meg, gramfort2014mne}. Specifically, we used a zero-phase finite impulse
response filter (FIR) with a Hamming window and with transition bands of 0.1Hz
and 10Hz for the low and high cut-off frequencies. The raw data was then segmented 100ms before word onset and 1s after
word onset ($t=0$ms corresponds to word onset). Finally, each resulting
segment was baseline-corrected between -100ms and 0ms, and decimated by 5 and
thus led a sampling frequency of 240Hz.
For each subject and each time sample relative to word onset, we
build an observation matrix $Y \in \mathbb{R}^{n \times d_y}$ of $n\approx$ 2700 words
by $d_y=301$ MEG channels (273 magnetometers and 28 compensation channels). Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

\subsubsection{Feature definition}
We annotated each word with three collinear features.
%
"Word Length' refers to the total number of letters, and is expected to specifically cause a variation in the early evoked MEG responses (i.e. from 100 ms after stimulus onset), because it variably engages different retinotopically-tuned visual areas in the primary visual cortex.
%
'Word Frequency', as derived with WordFreq \citep{speerwordfreq}, indexes how frequently each word appears in Dutch (using the Zipf logarithmic scale of \citep{van2014subtlex}). This feature is expected to specifically cause a variation in the late evoked MEG responses (i.e. from 400 ms), because it variably engages semantic processes in the temporal cortices \citep{kutas2011thirty}.
%
Finally, 'Word Function' indicates whether each word is a content word (i.e. a noun, a verb, an adjective or an adverb) or a function word (i.e. a preposition, a conjunction, a determinant, a pronoun or a numeral), and was derived from Spacy's part of speech tagger \citep{spacy2}. To our knowledge, this feature has not been thouroughly investigated with MEG. Its causal contribution to reading processes in the brain is therefore largely unknown.
%
To verify that B2B and other models would not inadequately identify non-causal features, we added a dummy feature, constructed from a noisy combination of "Word Length" and "Word Frequency":
\begin{equation}
  dummy = z(length) + z(frequency) + z(\mathcal{N})
\end{equation}
where $z$ transforms features into zero-mean and unit variance vectors.

This procedure yields an $X \in \mathbb{R}^{n \times d_x}$ matrix of $n\approx$ 2700 words by
$d_x=4$ features for each subject. Each of the columns of $X$ is normalized to
have a mean and a standard deviation of 0 and 1 respectively.

@@

We compare B2B to four standard models: forward regression, backward regression, CCA and PLS, as implemented in scikit-learn, and optimized with nested cross validation over 20 $\lambda$ l2 regularization parameters logarithmically spaced between $10^-4$ and $10^4$ or 1 to 4 caconical components.

We used the feature importance described in XXX algorithms XXX to assess the extent to which each feature $X_i$ specifically improves the prediction of held-out $Y$ data.

Corresponding results can be found in Figure \ref{fig:meg_twocurves}.

@@

%Forward regression was performed with an $\ell_2$-regularized least-squares regression $H \in \mathbb{R}^{d_x \times d_y}$ from $X$ to $Y$, and estimates the causal influence $E_{i,i}$ as $\| H_{i, :} \|^2$.
%Backward regression was performed with an $\ell_2$-regularized least-squares regression $G \in \mathbb{R}^{d_y \times d_x}$ from $Y$ to $X$, and estimates the causal influence $E_{i, i}$ as the Pearson correlation between the $i$-th column of $X$ and the $i$-th column of $YG$, using a validation sample. Finally, we used the back-to-back regression was performed using two $\ell_2$-regularized
%least squares regressions $G$ and $H$ as described in section \ref{sec:algorithm}, using a bagging factor of 10, and alpha cross-validated between $10^{-5}$ and $10^{5}$.

%The three types of analyses were implemented for each subject and each time
%sample independently. To estimate the reliability of these estimates in the
%population, we tested, using a one-sample $t$-test whether the distribution of
%the estimated contribution of each factor differed from baseline ($t=0$ms),
%as $\text{t-test}(\hat E_{t=0} - \hat E_{\tau})$, with $\tau$ ranging from
%150ms to 1s after word onset. We use the resulting t and p-values as proxy to
%quantify the performance of each model. These results are depicted in Figure
%3 of the Supplementary Material. For concision purposes, we only report this metric averaged
%between 150ms and 1s after word onset.

### 2019-09-24 paper/paper.tex
ICLR format.
wip paper

\usepackage{amsmath, amsfonts, microtype, xcolor, tikz, graphicx, hyperref, amsthm}
\usepackage[ruled, linesnumbered]{algorithm2e}
\usepackage[]{neurips_2019}

@@

\author{%
  Jean-Remi King\\
  CNRS\\
  \texttt{email} \\
  \And
  Fran\c{c}ois Charton\\
  Facebook AI\\
  \texttt{fcharton@fb.com}\\
  \And
  David Lopez-Paz\\
  Facebook AI\\
  \texttt{dlp@fb.com}
  \And
  Maxime Oquab\\
  Facebook AI\\
  \texttt{qas@fb.com}
}

@@

\begin{abstract}
Identifying causes from observations is at the core of science. This endeavor
is particularly challenging when i) potential factors are difficult to
manipulate individually and ii) observations are complex and multi-dimensional.
To address this issue, we introduce ``Back-to-Back'' regression (B2B), a
method designed to measure, from a set of co-varying factors, the causal
influences that most plausibly account for multidimensional observations. After
proving the consistency of B2B, we show that our method outperforms
least-squares and related regression and cross-decomposition techniques (e.g.
canonical correlation analysis and partial least squares) on two tasks: causal
identification and out-of-distribution prediction. Finally, we apply B2B to
neuroimaging recordings of 102 subjects reading word sequences. The results
show that the early and late brain responses caused by low- and high-level
word features respectively are more easily detected with B2B than with standard forward regression.
% conclude  open
\end{abstract}

@@

\section{Back to back regression}
\label{sec:algorithm}

@@

We consider the measurement of multivariate signal $Y$, generated from a set of variables $X$, via some unknown linear apparatus $F$.
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $E$, we denote by $XE$ the causal factors of $Y$.
%
In summary, our model can be formalized as:
%
\begin{equation}
    Y = (XE + N)F,\label{eq:model}
\end{equation}
%
where $N$ is a centered, unobserved noise variable.
%
While the triplet of variables $X$ and $N$ are independent, we allow each of them to have a general covariance matrix.
%
In practice, we observe $n$ samples $(X, Y)$ from the model.
%
This process, along with the sizes of all variables involved, is illustrated in Figure~\ref{fig:b2b}.
%
Given the model in Equation~\eqref{eq:model}, \textbf{the goal} of Back-to-Back Regression (B2B) is to estimate the matrix of causal influences $E$.

@@

Performing two regressions over the same data sample can result in overfitting, as spurious correlations in the data absorbed by the first regression will be leveraged by the second one.
%
To avoid this issue, we split our sample $(X, Y)$ at random into two equally-sized halves $(X_1, Y_1)$ and $(X_2, Y_2)$.
%
Then, the first regression is performed using $(X_1, Y_1)$, and the second regression is performed using $(X_2, Y_2)$.
%
To compensate for the reduction in sample size caused by the split, B2B is repeated over many random splits, and the final estimate $\hat E$ of the causal influence matrix is the average over the estimates associated to each split \citep{efron1992bootstrap}.
%
After obtaining $\hat{E}$, we can fit a final regression from $X \hat{E}$ to $Y$.

@@

\begin{algorithm}[H]
    %\SetAlgoLined
    \KwIn{input data $X \in \mathbb{R}^{n \times d_x}$, output data $Y \in \mathbb{R}^{n\times d_y}$, number of repetitions $m \in \mathbb{N}$.}
    \KwOut{estimate of causal influences $\hat{E} \in \mathbb{D}^{d_x \times d_x}$.}
    $\hat{E} \leftarrow 0$\;
    \For{$i = 1, \ldots, m$}{
        $(X, Y) \leftarrow \text{ShuffleRows}((X, Y))$\;
        $(X_1, Y_1), (X_2, Y_2) \leftarrow \text{SplitRowsInHalf}((X, Y))$\;
        $\hat{G} = \text{LinearRegression}(Y_1, X_1)$ \Comment*[r]{$\hat G = (Y_1^\top Y_1 + \Lambda_Y)^{-1} Y_1^\top X_1$}
        $\hat{H} = \text{LinearRegression}(X_2, Y_2 \hat{G})$ \Comment*[r]{$\hat H=(X_2^\top X_2 + \Lambda_X)^{-1} X_2^\top Y_2 \hat G$}
        $\hat{E} \leftarrow \hat{E} + \text{Diag}(\hat{H})$\;
    }
    $\hat{E} \leftarrow \hat{E} / m$\;
    $\hat{W} \leftarrow \text{LinearRegression}(X \hat{E}, Y)$\;
    \Return{$\hat{E}$, $\hat{W}$}
    \caption{Back-to-back regression.}
    \label{algorithm:b2br}
\end{algorithm}

@@

     Consider the B2B model from Equation $Y = (XE + N)F$, $N$ centred and full rank noise.
     %
     If $F$ and $X$ are full-rank on $Img(E)$, then, the solution of B2B, $\hat H$, will minimise
     %
     $\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$and satisfy $E\hat H = \hat H$
\end{theorem}
\begin{proof}

@@

The average value of the coefficients of $Diag(\hat H)$ is the trace of $\hat H$ divided by k. On average, since $X$ and $N$ are not correlated, $\Sigma_{X1}^{-1}\Sigma_{N1}$ should have a low condition numbers, and $\frac{Var(X)}{Var(X)+Var(N)}$ provide a good estimate of $\frac{1}{k}Tr(\hat H)$.

\subsection{Binarizing $\hat{E}$}

On the one hand, we have leveraged B2B to obtain the estimate $\hat{E}$, which is a diagonal matrix with real entries.
%
On the other hand, the true causal influence matrix $E$ is a binary matrix, which hard-selects the causal factors of $Y$ from $X$.
%
In this section, we provide one of the many possible recipes to binarize $\hat{E}$ and estimate the collection of causal factors.

In regression analysis, the traditional approach to this problem employs a $t$-test to check whether the regression coefficients differ from zero \citep{student1908probable}.
%
However, this test will not succeed here, since the estimate $\hat{E}$ is obtained from a double regression, and any employed regularization will add a bias to the diagonal of $\hat{E}$.
%
Instead, we treat the binarization of $\hat{E}$ as a clustering problem: separate the elements in the diagonal into a group of ``small values'', and a group of ``large values''.
%
More specifically, we propose to maximize the ratio of inter-group variance and to minimize the intra-group variance, over all possible splits of the diagonal into $p$ largest values and $d_x-p$ smallest values.
%
Letting $m_0$ and $m_1$ be the average values of the two clusters, $p$ and $d_x-p$ their size, and $v$ the total variance of the sample, we select the split maximizing the Sonquist-Morgan \citep{sonquist_morgan} criterion $\frac{p(d_x-p)}{d_x} \frac{(m_1 - m_0)^2}{v}$.
%
To binarize $\hat{E}$, set to one all the $p$ diagonal entries belonging to the ``large values'' group in the decided split, and setting to zero the rest of the $d_x-p$ diagonal entries.

@@

We perform two experiments to evaluate the performance of B2B: one on controlled synthetic data, and a second one on a real, large-scale magneto-encephalography dataset.
%
We use scikit-learn \citep{sklearn}.

@@

In the context of neuroimaging, the first regression of B2B relates to the first step of Representational Similarity Analysis (RSA) \citep{kriegeskorte2008representational}. Indeed, RSA analyzes the similarity of brain responses associated with specific conditions (e.g. distinct images), by (1) fitting one-against-all classifiers on each condition and (2) testing whether these classifiers can discriminate all other conditions. The resulting confusion matrix can then be analyzed in an unsupervised manner to reveal that some conditions lead to similar brain activity patterns. B2B differs from RSA in that (1) it uses regressions instead of classifications, and can thus generalize to new contexts and (2)
it is fully supervised.

More generally, B2B is related to CCA \citep{cca_hotelling}, but departs from it in four main aspects. First, CCA performs an additional eigenvector decomposition of what would be the $H$ coefficients in B2B \eqref{eq:solH}. Second, CCA does not typically optimize two distinct regularization parameters. Third, CCA does not use bagging between $G$ and $H$. Finally, CCA does not diagonalize nor binarize the $H$ matrix to estimate $\hat E$. More generally, CCA aims to find maximally correlated components between $X$ and $Y$, and does not assume any causal direction. By contrast, B2B assumes that $X$ causes $Y$. Together, these differences may explain why B2B can reliably outperform CCA on estimating causal influences and predicting $Y$ in novel environments (Fig. \ref{table:synthetic}).

@@

\bibliographystyle{plain}
\bibliography{paper}

>>>>

\usepackage{amsmath, amsfonts, microtype, xcolor, tikz, graphicx, hyperref, amsthm}
\usepackage[ruled, linesnumbered]{algorithm2e}
\usepackage{iclr2020_conference,times}

@@

\author{%
  Jean-Remi King\\
  CNRS\\
  \texttt{jeanremi.king@gmail.com} \\
  \And
  Fran\c{c}ois Charton\\
  Facebook AI\\
  \texttt{fcharton@fb.com}\\
  \And
  David Lopez-Paz\\
  Facebook AI\\
  \texttt{dlp@fb.com}
  \And
  Maxime Oquab\\
  Facebook AI\\
  \texttt{qas@fb.com}
}

@@

\begin{abstract}
Identifying causes from observations can be particularly challenging when i) potential factors are difficult to manipulate individually and ii) observations are complex and multi-dimensional.
To address this issue, we introduce ``Back-to-Back'' regression (B2B), a
method designed to efficiently measure, from a set of co-varying factors, the causal
influences that most plausibly account for multidimensional observations. After
proving the consistency of B2B and its links to other linear approaches, we show that our method outperforms
least-squares regression and cross-decomposition techniques (e.g.
canonical correlation analysis and partial least squares) on causal
identification. Finally, we apply B2B to
neuroimaging recordings of 102 subjects reading word sequences. The results
show that the early and late brain responses caused by low- and high-level
word features respectively are more reliably detected with B2B than with  other standard techniques.
% conclude  open
\end{abstract}

@@

\section{Back-to-Back regression}
\label{sec:algorithm}

@@

We consider the measurement of multivariate signal $Y$, generated from a set of variables $X$, via some unknown linear apparatus $F$.
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $E$, we denote by $XE$ the causal factors of $Y$.
%
In summary, our model can be formalized as:
%
\begin{equation}
    Y = (XE + N)F,\label{eq:model}
\end{equation}
%
where $N$ is a centered, unobserved noise variable.
%
While the triplet of variables $X$ and $N$ are independent, we allow each of them to have a general covariance matrix.
% what is a general covariance matrix?
%
In practice, we observe $n$ samples $(X, Y)$ from the model.
%
This process, along with the sizes of all variables involved, is illustrated in Figure~\ref{fig:b2b}.
%
Given the model in Equation~\eqref{eq:model}, \textbf{the goal} of Back-to-Back Regression (B2B) is to estimate the matrix of causal influences $E$.

@@

Performing two regressions over the same data sample can result in overfitting, as spurious correlations in the data absorbed by the first regression will be leveraged by the second one.
%
To avoid this issue, we split our sample $(X, Y)$ at random into two subsets $(X_1, Y_1)$ and $(X_2, Y_2)$.
%
Then, the first regression is performed using $(X_1, Y_1)$, and the second regression is performed using $(X_2, Y_2)$.
%
To compensate for the reduction in sample size caused by the split, B2B is repeated over many random splits, and the final estimate $\hat E$ of the causal influence matrix is the average over the estimates associated to each split \citep{efron1992bootstrap}.
%
To accelerate this ensembling procedure, we implemented an efficient leave-one-out cross-validation scheme as detailed in \citep{RifkinLippert2007} % http://cbcl.mit.edu/publications/ps/MIT-CSAIL-TR-2007-025.pdf
, as follow:
%

\begin{equation}
\hat{Y} = (\Sigma_X G Y - diag(\Sigma_X G) Y) / diag(I - \Sigma_X G)
\end{equation}

where $\Sigma_X$ is the $X$ kernel matrix and where $G$ is computed with an eigen decomposition of $X$:
\begin{equation}
  \begin{aligned}
  \Sigma_X = Q V Q^T \\
  G = Q (V + \alpha*I)^{-1} Q^T\\
  \end{aligned}
\end{equation}
where $Q$, $V$ and $\alpha$ are the eigen vectors, eigen values and regularization, respectively.

%
% After obtaining $\hat{E}$, we can fit a final regression from $X \hat{E}$ to $Y$.

@@

% JR: shall we update algorithm with LOO trick?
\begin{algorithm}[H]
    %\SetAlgoLined
    \KwIn{input data $X \in \mathbb{R}^{n \times d_x}$, output data $Y \in \mathbb{R}^{n\times d_y}$, number of repetitions $m \in \mathbb{N}$.}
    \KwOut{estimate of causal influences $\hat{E} \in \mathbb{D}^{d_x \times d_x}$.}
    $\hat{E} \leftarrow 0$\;
    \For{$i = 1, \ldots, m$}{
        $(X, Y) \leftarrow \text{ShuffleRows}((X, Y))$\;
        $(X_1, Y_1), (X_2, Y_2) \leftarrow \text{SplitRowsInHalf}((X, Y))$\;
        $\hat{G} = \text{LinearRegression}(Y_1, X_1)$ \Comment*[r]{$\hat G = (Y_1^\top Y_1 + \Lambda_Y)^{-1} Y_1^\top X_1$}
        $\hat{H} = \text{LinearRegression}(X_2, Y_2 \hat{G})$ \Comment*[r]{$\hat H=(X_2^\top X_2 + \Lambda_X)^{-1} X_2^\top Y_2 \hat G$}
        $\hat{E} \leftarrow \hat{E} + \text{Diag}(\hat{H})$\;
    }
    $\hat{E} \leftarrow \hat{E} / m$\;
    $\hat{W} \leftarrow \text{LinearRegression}(X \hat{E}, Y)$\;
    \Return{$\hat{E}$, $\hat{W}$}
    \caption{Back-to-back regression.}
    \label{algorithm:b2br}
\end{algorithm}

@@

     Consider the B2B model from Equation $Y = (XE + N)F$, $N$ centred and full rank noise.
     %
     If $F$ and $X$ are full-rank on $Img(E)$, then, the solution of B2B, $\hat H$, will minimize
     %
     $\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$and satisfy $E\hat H = \hat H$
\end{theorem}
\begin{proof}

@@

The average value of the coefficients of $Diag(\hat H)$ is the trace of $\hat H$ divided by k. On average, since $X$ and $N$ are not correlated by definition, $\Sigma_{X1}^{-1}\Sigma_{N1}$ should have a low condition numbers, and $\frac{Var(X)}{Var(X)+Var(N)}$ provide a good estimate of $\frac{1}{k}Tr(\hat H)$.

@@

We perform two experiments to evaluate B2B: one on controlled synthetic data, and a second one on a real, large-scale magneto-encephalography dataset.
%
We use scikit-learn's CCA, PLS, and RidgeCV objects to compare B2B against the standard baselines \citep{sklearn}.

@@

In the context of neuroimaging, the first regression of B2B relates to the first step of Representational Similarity Analysis (RSA) \citep{kriegeskorte2008representational}. Indeed, RSA analyzes the similarity of brain responses associated with specific categorical conditions (e.g. distinct images), by (1) fitting one-against-all classifiers on each condition and (2) testing whether these classifiers can discriminate all other conditions. The resulting confusion matrix can then be analyzed in an unsupervised manner to reveal that some conditions lead to similar brain activity patterns. B2B differs from RSA in that (1) it uses regressions instead of classifications, and can thus generalize to new contexts and (2)
it is fully supervised.

% THIS NEEDS TO BE UPDATED: see de chevigner's latest paper
More generally, B2B is related to CCA \citep{cca_hotelling}, but departs from it in four main aspects. First, CCA performs an additional eigenvector decomposition of what would be the $H$ coefficients in B2B \eqref{eq:solH}. Second, CCA does not typically optimize two distinct regularization parameters. Third, CCA does not use bagging between $G$ and $H$. Finally, CCA does not diagonalize nor binarize the $H$ matrix to estimate $\hat E$. More generally, CCA aims to find maximally correlated components between $X$ and $Y$, and does not assume any causal direction. By contrast, B2B assumes that $X$ causes $Y$. Together, these differences may explain why B2B can reliably outperform CCA on estimating causal influences and predicting $Y$ in novel environments (Fig. \ref{table:synthetic}).

@@

\bibliographystyle{iclr2020_conference}
\bibliography{paper}
